### 2019-10-27 tdabench-paper/tdabench-paper.Rmd
Added tentative figure descriptions

---
title: Comparing efficiency of R packages implementing calculation of persistent homology
author:
  - name: Author One
    affiliation: Affiliation
    address:
    - line 1
    - line 2
    email:  author1@work
  - name: Author Two
    affiliation: Affiliation
    address:
    - line 1
    - line 2
    email:  author2@work
abstract: >
  An abstract of less than 150 words.
output:
  rticles::rjournal_article:
    includes:
      in_header: preamble.tex
---

## Introduction

Introductory section which may include references in parentheses
\citep{R}, or cite a reference such as \citet{R} in the text.

## Methods

This section may contain a figure such as Figure \ref{figure:rlogo}.

\begin{figure}[htbp]
  \centering
  \includegraphics{Rlogo}
  \caption{The logo of R.}
  \label{figure:rlogo}
\end{figure}

## Results

Figure 1: Comparison of persistent homology calculation runtimes across R packages for a torus. Panel 2. Horizontal axis: number of points in torus. Vertical axis: runtime. Point chart. Point color: underlying engine. Point shape: same as color. Panel 1. Torus showing users visual representation of dataset/point cloud. Goal: overview showing speed differences across engines on a canonical shape.

Figure 2: Comparison of persistent homology calculation runtimes across R packages for n-dimensional boxes (n = 2, 3, 4, 5). For each dimension, features up to n-1 dimensions are calculated. Panel 1. Visual representation of 2-d box (maybe points, maybe box, maybe scattered both). Panel 2. Visual representation of 3-d box (maybe points, maybe box, maybe both scattered). Panels 3-6 (differing n={2,3,4,5} for dimension). Point shape/color: underlying engine. Horizontal axis: number of points. Vertical axis: runtime.

Figure 3: Runtime of persistent homology calculation as a function of underlying engine (4d noisy circle/annulus). Panel 1: visualize noisy circle - use imagination for 4d. Panel 2. Horizontal axis: number of points. Vertical axis: runtime. Point color/shape: underlying engine. Goal: show that TDAstats and GUDHIalpha are sufficiently faster than GUDHI and Dionysus, so the latter pair can be ignored for rest of figures.

Figure 4: Runtime of persistent homology calculation as a function of underlying engine and feature dimension (4d noisy circle/annulus). Panel 1 = GUDHI alpha engine. Horizontal axis = number of points. Vertical axis = runtime. Point color/shape = feature dimension. Panel 2 = panel 1, but with TDAstats engine. Goal: show how increasing dimension slows down the calculation time.

Figure 5: Memory use of Rips complex vs alpha complex. Engine = GUDHI(alpha). Panel 1 = 4-sphere. Panel 2 = 3-annulus. Panel 3 = torus. Panel 4 = 5-box. For each panel: point color/shape = engine (GUDHI vs alpha); horizontal axis = number of points; vertical axis = memory use; feature dimension = data dimension - 1 for all panels. Goal: compare memory
POINT OUT THAT RIPSER DOESN'T MEASURE MEMORY B/C NO BOUNDARY MATRIX, BUT LIKELY SOMEWHERE IN BETWEEN THE TWO

Figure 6: Same as Figure 5, but TDAstats vs GUDHIalpha for runtime.

## Discussion

This file is only a basic article template. For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](https://journal.r-project.org/share/author-guide.pdf).
\bibliography{RJreferences}